---
title: "Bank Marketing Success"
author: "Demetri Lee"
date: "April 1, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(tidyverse)
library(summarytools)

#bankfull <- read_delim(file = "bank-additional.csv", delim = ";", col_types = "iccccccccciiiicdddddc")
bankfull <- read_delim(file = "bank-additional.csv", delim = ";", col_types = "ifffffffffiiiifdddddf")

colnames(bankfull) <- c("age", "job", "marital", "edu_lvl", "cred_default", "mortgage", "loan",
                        "contact_typ", "month", "day_of_week", "duration_sec", 
                        "contact_cnt", "days_passed", "prev_contact_cnt", "prev_outcome",
                        "emp_var_rate", "cons_price_idx", "cons_conf_idx", "euribor3m", "num_employed",
                        "subscribed")
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

A Portuguese bank uses telemarketing campaigns to engage their customers. 

Each customer engagement has been recorded for evaluation. 

Such data can be used to extract insights for making better campaign decisions.


## Motivation

The Bank essentially has a binary classification problem, whether a customer 
suscribes or not. 

Marketing campaigns generally do not discriminate how customers are targeted. 

Therefore, some campaigns are irrelevant to some customers and needless 
resources may be spent on rejections. 


## Data

Interactions with over 41,000 clients were recorded from May 2008 to Nov 2010.

Each record contains 21 different attributes for evaluation.

The attribute of interest, `subscribed`, will indicate a successful 
subscription from a client.

Data was provided by Sérgio Moro (ISCTE-IUL), Paulo Cortez (Univ. Minho) and 
Paulo Rita (ISCTE-IUL).

## Exploratory Data Analysis

At first glance there is an overwhelming amount of clients who choose not to 
subscribe to a term deposit. This means the data set is imbalanced.

```{r subscribed}
ggplot(bankfull, aes(x = subscribed, fill = subscribed)) + geom_bar() + 
  xlab("subscribed") + 
  ylab("count")
```


## Exploratory Data Analysis

But looking at subscriptions from a previous campaign tells us more than 
half the subscribers will re-subscribe.

```{r prev_outcome}
bankfull_prev <- filter(bankfull, prev_outcome != "nonexistent")
ggplot(bankfull_prev, aes(x=prev_outcome, fill=subscribed)) + 
  geom_bar(position="fill") + 
  xlab("previous campaign") + 
  ylab("ratio")
```


## Contact Type

It was also observed contacting clients via cellular phone yields at least 2 
times more success than contacting a client through their landline telephone.

```{r contact_typ}
ggplot(bankfull, aes(x=contact_typ, fill=subscribed)) + 
  geom_bar(position="fill") + 
  xlab("method of contact") + 
  ylab("ratio")
```


## Machine Learning

The data was partitioned into a training, validation, and test set.

1. Two predictive models were created from client responses in the training set.

2. The same two models were used to predict client responses in the validation set.

3. Finally the optimal model from the prior step was used with the test set to 
  estimate future results. 

## Predictive Modeling

Since this is a binary classification problem, the following two Machine 
Learning models were used to predict subscriptions:

* Logistic Regression

* Random Forest

It was noted the data set is extremely imbalanced as there is overwhelming 
amount of clients that do not subscribe. Hence the statistical Kappa value is 
more helpful than using accuracy to determine which model will give us the 
best results.


## Evaluating Model Performance

The Kappa value for the Random Forest model is higher than the Logistic 
Regression model so the former will be the model of choice for the test data.

Since there is a high volume of rejection, the accuracy of a prediction for 
rejection is overwhelmingly high. Yet we are interested in successful 
subscriptions, which are true positive results.


## Recommendations

1. Reach out to subscribers from previous campaign as more than 50% of 
  subscribers will re-subscribe.

2. Contact clients via their cellular phone as they are more than twice as 
  likely to subscribe compared to clients contacted via a regular telephone.

3. The Random Forest model is a better predictor for this bank marketing 
  dataset. However, there aren't any tangible results that can be applied. 
  For better explanantory analysis I would recommend using the results from the Logistic 
  Regression model.
